---
title: "Experiment 1 Data Analysis"
output:
  html_document:
    df_print: paged
---

Loading libraries and functions for data import:

```{r}

library(plyr)
library(lme4)
library(lmerTest)
library(ggplot2)
library(tidyverse)
library(broom)
library(emmeans)
library(ordinal)
library(brms)


results <- read.csv("https://159d3d98-11fc-4642-8e29-43c994fa07d0.usrfiles.com/ugd/159d3d_ab22ce000af244c4a2410dc21d6c6a68.csv")
view(results)
table(results$Condition)


```

First overview plots:

```{r}

results$Condition <- (as.factor(as.character(results$Condition)))

levels(results$Condition) <- c("NoPsFirst", "NoPsSecond", "PsFirst", "PsSecond", "EISimplePs", "SSimplePs")

results <- as_tibble(results)

results <- results %>%
          mutate(Condition = fct_relevel(Condition, 
             c("PsFirst", "PsSecond", "NoPsFirst", "NoPsSecond", "EISimplePs", "SSimplePs")))

resultsPlot <- results %>%
  group_by(Condition) %>%
  dplyr::summarize(n = n(), MeanValue = mean(Value), sd = sd(Value))%>%
  mutate( se=sd/sqrt(n))  %>%
  mutate( ci=se * qt((1-0.05)/2 + .5, n-1))

colours <- c("#727272", "#f1595f", "#79c36a", "#599ad3", "#f9a65a" , "#9e66ab")

ggplot(resultsPlot, aes(x = Condition, y = MeanValue, fill=Condition))+
  geom_bar(stat = "identity") +
  ylim("1", "2", "3", "4", "5", "6", "7")+
  geom_errorbar(aes(ymin=MeanValue-se, ymax=MeanValue+se), width=.1) +
  scale_fill_manual(values=colours)+
  theme(axis.text.x = element_text(angle = 45)) +
  theme_classic() +
  labs(title = "Mean rating per condition",
       y = "Mean rating",
       x = "Condition")
        

```

Setting up factors for analyses:

```{r}

results$Condition <- as_factor(results$Condition)
results$ID <- as_factor(results$ID)
results$Trigger <- as_factor(results$Trigger)
results$Value <- as_factor(results$Value)

results <- results %>%
      mutate(PsType = fct_collapse(Condition,
        Ps = c("SSimplePs", "EISimplePs", "PsFirst", "PsSecond"),
        NoPs = c("NoPsFirst", "NoPsSecond")
      )) %>%
      mutate(Order = fct_collapse(Condition,
        First = c("PsFirst", "NoPsFirst"),
        Second = c("PsSecond", "NoPsSecond"),
      ))%>%
      mutate(Order = fct_relevel(Order, c("First", "Second"))) %>%
      mutate(SupType = fct_collapse(Condition,
        NoPriorSupp = c("PsFirst", "EISimplePs"),
        PriorSupp = c("PsSecond", "SSimplePs")
      ))%>%
      mutate(CompType = fct_collapse(Condition,
                    Simple = c("SSimplePs", "EISimplePs"),
                    Complex = c("PsFirst", "PsSecond", "NoPsFirst", "NoPsSecond")
                                      )) %>%
      mutate(SupType = fct_relevel(SupType, c("PriorSupp", "NoPriorSupp"))) %>%
      mutate(OrderNum = as.numeric(Order)-1.5) %>%       ##### here and below, using this to numerically force the first two levels of the factor to be -0.5 and 0.5; any other levels will be removed via filter() in the relevant 2x2 analyses, so no harm done.
      mutate(PsTypeNum = as.numeric(PsType)-1.5) %>%
      mutate(SupTypeNum = as.numeric(SupType)-1.5) %>%
      mutate(CompTypeNum = as.numeric(CompType)-1.5)

filter(results, Order %in% c("First", "Second"))
filter(results, SupType %in% c("PriorSupp", "NoPriorSupp"))



```

Check difference between EISimplePs and SSimplePs:

```{r}

EIvSSimple.ordinal <- clmm(Value ~ Condition + (1|ID) + (1+Condition|Trigger), filter(results, Condition %in% c("EISimplePs", "SSimplePs")), link = "logit")

summary(EIvSSimple.ordinal)
```


Testing for Ps-NoPs interaction (predicted by asymmetry-based accounts):

```{r}

PsNoPs.ordinal <- clmm(Value ~ PsTypeNum*OrderNum + (1|ID) + (1+PsTypeNum*OrderNum|Trigger), filter(results, Order %in% c("First", "Second")), link = "logit")

summary(PsNoPs.ordinal)

PsNoPs.ordinal2 <- clmm(Value ~ PsTypeNum*OrderNum + (1|ID) + (1+PsTypeNum+OrderNum|Trigger), filter(results, Order %in% c("First", "Second")), link = "logit")


anova(PsNoPs.ordinal,PsNoPs.ordinal2)


PsNoPs.ordinal3 <- clmm(Value ~ PsTypeNum*OrderNum + (1|ID) + (1+PsTypeNum|Trigger), filter(results, Order %in% c("First", "Second")), link = "logit")

anova(PsNoPs.ordinal,PsNoPs.ordinal3)

PsNoPs.ordinal4 <- clmm(Value ~ PsTypeNum*OrderNum + (1|ID) + (1+OrderNum|Trigger), filter(results, Order %in% c("First", "Second")), link = "logit")

anova(PsNoPs.ordinal,PsNoPs.ordinal4)

```

Final model is PsNoPs.ordinal3

```{r}
summary(PsNoPs.ordinal3)
```

No interaction, no significant condition effects. Note: OrderNum slope for Trigger also converges, but accounts for much less variance; signifance pattern doesn't change.

The Bayesian part of the analysis:

```{r}
head(filter(results, Order %in% c("First", "Second")))
results$Value <- as.numeric(as.character(results$Value)) #make Value a factor (model below is cumulative and requires it)
# Priors from Mandelkern et al. brm analysis (newly conducted; no Bayesian stats in original)
priors <- c(
    prior(normal(-2.26, 0.22), class = "Intercept",
        coef = "1"),
  prior(normal(-1.18, 0.21), class = "Intercept",
        coef = "2"),
  prior(normal(-0.52, 0.21), class = "Intercept",
        coef = "3"),
  prior(normal(-0.06, 0.21), class = "Intercept",
        coef = "4"),
  prior(normal(0.79, 0.21), class = "Intercept",
        coef = "5"),
  prior(normal(2.17, 0.22), class = "Intercept",
        coef = "6"),
  prior(normal(0.79, 0.32), class = "b", 
        coef = "PsTypeNum"),
  prior(normal(1.04, 0.22), class = "b",
        coef = "OrderNum"),
  prior(normal(-1.49, 0.38), class = "b",
        coef = "PsTypeNum:OrderNum")
)


PsNoPs.brm <- brm(Value ~ PsTypeNum*OrderNum + (1|ID) + (1+PsTypeNum|Trigger),
                 family=cumulative("logit"),  
                 control = list(adapt_delta = 0.98),
                 save_pars = save_pars(all = TRUE),
                 iter = 20000,
                 prior =  priors,
                 filter(results, Order %in% c("First", "Second")), backend = "cmdstanr")

PsNoPs.brm

PsNoPs.brmStan <- brm(Value ~ PsTypeNum*OrderNum + (1|ID) + (1+PsTypeNum|Trigger),
                 family=cumulative("logit"),  
                 control = list(adapt_delta = 0.98),
                 save_pars = save_pars(all = TRUE),
                 iter = 0,
                 chains = 0,
                 prior =  priors,
                 filter(results, Order %in% c("First", "Second")), backend = "rstan")

priors <- c(
    prior(normal(-2.26, 0.22), class = "Intercept",
        coef = "1"),
  prior(normal(-1.18, 0.21), class = "Intercept",
        coef = "2"),
  prior(normal(-0.52, 0.21), class = "Intercept",
        coef = "3"),
  prior(normal(-0.06, 0.21), class = "Intercept",
        coef = "4"),
  prior(normal(0.79, 0.21), class = "Intercept",
        coef = "5"),
  prior(normal(2.17, 0.22), class = "Intercept",
        coef = "6"),
  prior(normal(0.79, 0.32), class = "b", 
        coef = "PsTypeNum"),
  prior(normal(1.04, 0.22), class = "b",
        coef = "OrderNum")
)


PsNoPsNullInt.brm <- brm(Value ~ PsTypeNum+OrderNum + (1|ID) + (1+PsTypeNum|Trigger),
                 family=cumulative("logit"),  
                 control = list(adapt_delta = 0.98), 
                 save_pars = save_pars(all = TRUE),
                 iter = 20000,
                 prior =  priors,                 
                 filter(results, Order %in% c("First", "Second")))



margLogLik_linear <- bridge_sampler(PsNoPs.brm,  silent = TRUE)

margLogLik_null <- bridge_sampler(PsNoPsNullInt.brm, silent = TRUE)

(BF_ln <- bayes_factor(margLogLik_linear, margLogLik_null))

```


Testing for condition differences relative to Support Control:

```{r}

PsNoPs.emm = emmeans(PsNoPs.ordinal3, specs = pairwise ~ OrderNum|PsTypeNum)

PsNoPs.emm$contrasts %>%
  rbind()

```

No significant effects of Order in either Ps or No-Ps

Next: testing interaction between CompType and SupType:

```{r}

Support.ordinal <- clmm(Value ~ CompTypeNum * SupTypeNum + (1|ID) + (1+CompTypeNum*SupTypeNum|Trigger), filter(results, PsType =="Ps"), link = "logit")

summary(Support.ordinal)

Support.ordinal2 <- clmm(Value ~ CompTypeNum * SupTypeNum + (1|ID) + (1+CompTypeNum+SupTypeNum|Trigger), filter(results, PsType =="Ps"), link = "logit")

Support.ordinal3 <- clmm(Value ~ CompTypeNum * SupTypeNum + (1|ID) + (1+SupTypeNum|Trigger), filter(results, PsType =="Ps"), link = "logit")

Support.ordinal4 <- clmm(Value ~ CompTypeNum * SupTypeNum + (1|ID) + (1+CompTypeNum|Trigger), filter(results, PsType =="Ps"), link = "logit")

anova(Support.ordinal,Support.ordinal2)
anova(Support.ordinal,Support.ordinal3)
anova(Support.ordinal,Support.ordinal4)
```

Including by-trigger random slope for interaction doesn't significantly improve model fit. Including by-trigger random slope for CompType does make a difference. Including by-trigger random slope for SupType doesn't make a difference. So, our final model is Support.ordinal4:

```{r}
summary(Support.ordinal4)

```

Significant interaction (plus main effect dominated by interaction)

Follow-up pairwise comparisons:

```{r}
Support.emm <- emmeans(Support.ordinal4, specs = pairwise ~ SupTypeNum|CompTypeNum)



Support.emm$contrasts %>%
  rbind()


```

Support-Ps-Cond is rated significantly higher than EI-Ps-Cond. The Disj conditions are NOT significantly lower in rating than S-Ps-Cond.

